Global adjoint tomography -- Model GLAD-M25

### 2019-01-23 main.tex
update window plots

With the development of high performance computing and numerical methods,
it is now possible to using adjoint method to image the interior the globe
using thousands of earthquakes.

There are few agreements about the earth model and yet many unknown to be
resolved. There are quite a lot of global models exits however there is no such con-sense.

The GLAD-M25 is the first earth model that incorporate the P and S wave speed using full waveform inversion. Also, people are using specific set of phases in their inversion. For P models, people are using P phase and other body wave phases and for S models, people are using mostly S wave phases and surface waves to constrain the structure. Those two kinds of models usually reveals very different parts of the earth. For example, for P wave, people used it the image the subduction slabs and it is not very good at image lower mantle structures, such as plumes and lower mantle convection. For S waves, it is the verse verso.

@@

Here, we present our GLAD-M25 earth model, with significantly improvements on the plumes and subductions.

@@

@@

In summary, the new model performs well for the held-out dataset,
both in terms of misfit reductions and measurement distributions
in all twelve measurement categories. We expect to see similar
behavior for any future earthquakes.

Seismology is driven by data. Back into time, seismologists would look at
single seismograms and pick valuable information out the them. Now, we have
ten thousands of seismograms deployed globally and it will be a pity not
to take all the information we have. Even though GLAD-M25 is following-up
of GLAD-M15. It has many superior advantages and improvements. We increase
the dataset from 253 to 1480 earthquakes, with carefully inverted sources
in the 3D earth structures. We have 360 earthquakes used in the test
dataset, which is already larger than the most of the dataset other people used
for their inversion. We culled every seismograms from various data centers,
thanks to the open environment and data sharing spirit of our seismology community.
With very careful checks by assessing the integrity of instrument response
and deployment, evaluating the signal-to-noise ratios, picking the time windows
using over 20 criteria, cleaning windows by their statistical behavior, and
calculating weights to balance the uneven distribution of earthquakes and
seismic stations, we really have done a lot to ensure the cleanness of the
data assimilated into our inversion and make sure every piece of value information
will be used but rejecting the crap data.

There are also quite a lot of technique challenges on the way but we are seeking
the most cutting-edge technologies to solve them. ADIOS was introduced to save mesh
and kernel related data, since we are saving petabytes of wavefiled snapshots
in a few hours of simulation and it is critical even for the most advanced
supercomputers in the world. ASDF is developed by us to store the seismograms
since we have ten millions of time series data from thousands of earthquakes.
We have to get access and process them in a very efficient manner that most of
the existing data format just won't meet our requirements. To handle the complex
inversion workflow, we are working RADICAL group to develop the workflow management
tools that will help manage every piece of task and make the user aware if there
is anything bad happened during the computation. Those tools that help the
automation of the workflow, is the "guard" that stand there and monitor things
to make sure our scientific results away from contamination. Besides, the
developer of SPECFEM packages, they dedicate more than ten years of work,
to develop the software and make it fast, efficient and robust for global scale of simulation.
They are working closely with hardware vendors such as Cray, IBM and NVidia to adapt
the code to architecture of the most advanced supercomputers. We are soon
moving our simulation to Summit, the world No.1 supercomputer at the Oak Ridge
National Lab.

\newpage
\begin{acknowledgments}
%We want to give thanks to the Oak Ridge National Lab, for providing
%  excellent support and computational resource that facilitate our research.
%  We want to thanks IRIS and ORFEUS for providing seismic datas used in this
%  study. We want to acknowledge Hejun Zhu, Min Chen, Carl Tape and Qinya Liu,
%  for providing usefull tools and great suggetions these years.
%  We want to thanks Ebru Bozdag, for providing GLAD-M15, data and software
%  she used in the first generation model.

>>>>

In the last few decades, we have witnessed great improvments in the global
seismic imaging. There are different types of earth models published using
various data types and methodologies\citep{}. The theory of Finite-frequcy
was brought up in the early 1980\citep{}, overcomes very limitations
in the ray-based theory. However, it is still the very recent
decade that people are able to use it in the global scale inverison, 
thanks to the advance in the high-performance computing and numerical methods.

There are gaps between the global and regional models. Regional models, tends to
resolve structures at shorter period bands such as 10 - 15 seconds\citep{},
but lacks of resolution in depth. 
On the other hand, global models, resolved structures as deep as 
the core-mantle boundary, but only starting from ~60 seconds period bands
at most, mainly due to the high computational cost\citep{}. Recent years
we have seen models moves to 30 seconds in waveform-type inversion but
there is still a clear gap between the global and regional inversions.
Owing to the recent advances in high-performance computing and numerical
methods, it is possible to simulate seismic waves in an unprecendental
resolution at global scale, to close the gap between regional and global
studies.

In our previous model paper\citep{}, advantanges of spetral-element methods
and adjoint tomography has been thoroughly discussed. Crust and discontinuity
is honored in our mesh. Thus, there is no crust correction needed in our
inversion procedures. Adjoint methods, can fully take the 3D earth structure
an attenuation into account, for both forward simulation and adjoint simulation,
calculating the frechet direvative in a more precise way.
However, when pushing for high frequecy, it soon becomes very
demanding on computational resourse.

Other than the simulation part, the data assimilation has been another key factor
to the success of tomography. Ray-based travel-time inversion has accumulated
huge amount of data along it way. However, they usually focused on certain type of phases.
relying on both open and customized travel time database.
Models, such as GAP-P4 and UU-P07, used millions of travel time measurements, but only
used P related phases. 
The finite frequcy methods is developed recently so there are less consense of the dataset
used in the research. On the other hand, the methods also gives more freedom when it comes
to data assimilation. It then become a critial question for which part of the data should
be assimilated into the inverison. 
Recent years have seen a lot of good quality seismic networks deployed, such as US array.
However, there is few global model that take the advantage of network yet. It would be
a pity not to utlized it, but data assimilation should be handled very carefully since it
will directly impact the image in a few ways. First, it is the data quality. only good
seismograms and phases should be including into the inversion. Second, the high density
distribution of regional networks would skew the inversion so proper weightings should be
taken care of.

@@

In the following sections, we will present the GLAD-M25 earth model in details.
From section 2 - 4, we will focus on the data, including the starting model,
earthquakes and seismic data used in this study.  In section 5 and 6, we will
discuss the model parametrization and the choice of misfit function we made.
Section 7 and 8 describes the adjoint tomography workflow and the evolution
of inversion, and the improvments we made with the iterations. Section 9 will
focus on model evaluation part, which we used a few techniques to assess the
quality of inverison. Section 10 will address on presenting images of our model,
together with whole bunch of other models, global and regional. We conclude in 
section 10, summarizing our model and diretion of future works.

@@

\section{Discussion}

Here, we presented our GLAD-M25 models. Starting from GLAD-M25, we
conducted 10 more iterations, with significantly more earthquakes
(up to 1480 earthquakes)and seismic data(up to 18.2 million time
windows) added in the inversion in a progressive way. It is the first
global transerse isotropoic model that brings the resolution down to
17 secs, but also invert the $V_p$ and $V_s$ indenpendently. Due to the
nature of spetral-element methos, our model meshes the crust and mantal
, without introducing the "crustal correction". Our solver also enables
us the take the full-attenuation into accout, both in forward and adjoint
simulations, using undo attenuation at the cost of more computation
hours and disk space.

GLAD-M25 is evaluated in many ways. First, it shows significant misfit
reduction in all measurements categories, but also in the variance
reduction in travel-time measurements. For the 360 held-out earthquakes,
the behavior is similar to what we have in our inversion dataset. We
presented our global $V_s$ and $V_p$, together with other well-known global
models and at large scale they agree with each other. However, when we focus
on regional studies and compare with regional models, our model shows superior
more detailed structures compared to other global models, staying at the 
same resolution of regional models. Our models shows great resolution in both
plumes and suduction zones, using the independently inverted $V_p$ and $V_s$
velocity models.

@@

Seismology is driven by data. Back into time, seismologists would look at
single seismograms and pick valuable information out the them. Now, we have
ten thousands of seismograms deployed globally and it will be a pity not
to take all the information we have. Even though GLAD-M25 is following-up
of GLAD-M15, It has many superior advantages and improvements. We increase
the dataset from 253 to 1480 earthquakes, with carefully inverted sources
in the 3D earth structures. We have 360 earthquakes used in the test
dataset, which is already larger than the most of the dataset other people used
for their inversion. We culled seismograms from various data centers,
thanks to the open environment and data sharing spirit of our seismology community.
With very careful checks by assessing the integrity of instrument response
and deployment, evaluating the signal-to-noise ratios, picking the time windows
using over 20 criteria, cleaning windows by their statistical behavior, and
calculating weights to balance the uneven distribution of earthquakes and
seismic stations, we really have done a lot to ensure the cleanness of the
data assimilated into our inversion and make sure every piece of value information
will be used but rejecting crap data.

However, global tomography research has progressed to a stage that not only relying
on the data and methodology, but also involving a lot of technique challenges
under the hood.
It has driven us to seek for the most cutting-edge technologies to solve them.
ADIOS was introduced to save mesh and kernel related data, since we are saving
petabytes of wavefiled snapshots in a few hours of simulation and it is critical
even for the most advanced supercomputers in the world.
ASDF is developed by us to store the seismograms since we have ten millions of
time series data from thousands of earthquakes. We have to get access and process
them in a very efficient manner that most of the existing data format just won't
meet our requirements.
To handle the complex inversion workflow, we are working RADICAL group to develop
the workflow management tools that will help manage every piece of task and make
the user aware if there is anything bad happened during the computation. Those tools that help the
automation of the workflow, is the "guard" that stand there and monitor things
to make sure our scientific results away from contamination. Besides, the
developer of SPECFEM packages, they dedicate more than ten years of work,
to develop the software and make it fast, efficient and robust for global scale of simulation.
They are working closely with hardware vendors such as Cray, IBM and NVidia to adapt
the code to architecture of the most advanced supercomputers. We are soon
moving our simulation to Summit, the world No.1 supercomputer at the Oak Ridge
National Lab.

For the future work, inversion for azimuthal anisotropy and attenuation is on the way.
The methodolody has already be demonstrated in a few regional studies, such as
\citep{}. We will start with our current dataset but increase our earthquakes to
more than $6,000$ in a progressive way , which will covered most of the
earthquakes in Global CMT project. We also plan to push our model resolution to
9 secs, with the Summit supercomputer available at Oak Ridge national lab. It will be
unprecedental in sesimology that pushes global tomography to such resolution, meanwhile
it will pose new chanllenges. Ensuring the reliability and stability on such a large machine
and associated file system, for example, will be one of those challenges. Even though the
computing powers has grown significantly, the storage has not. Our 17 sec has already
generated a few petabytes of wavefields files and 9 sec will generate at least one order
of magnitudes more files.

There are certains things that we would like to address on in the near future. First is the
the seismic data distribtuion. As we know, most of the seismic stations are distributed on
the continents, while the oceans are most left blank, with only a few stations on a few
islands. Floating arrays\citep{} might be promising in the future but it all comes down to the 
data processing and data quality. Second, harnessing such a large data set is difficult.
Artifical intelligence may be introduced and help us to make decisions when assimilating
measurements\citep{}.

\begin{acknowledgments}
%We want to give thanks to the Oak Ridge National Lab, for providing
%  excellent support and computational resource that facilitate our research.
%  We want to thanks IRIS and ORFEUS for providing seismic datas used in this
%  study. We want to acknowledge Hejun Zhu, Min Chen, Carl Tape and Qinya Liu,
%  for providing usefull tools and great suggetions these years.
%  We want to thanks Ebru Bozdag, for providing GLAD-M15, data and software
%  she used in the first generation model.
